---
title: "Final Project"
author: "Yidan Zhu"
date: "3/2/2022"
output:pdf_document: default
---

```{r}
df<-read.csv('diabetes.csv')
```

#Main Question of Interest: Influence of number of pregancies on Diabetes
#Main Variable of Interest: Pregnancies
#Confounder Variables: BloodPressure, SkinThickness, BMI, Age, and DiabetesPedigreeFunction
#Response: Outcome

#Excluding Glucose and Insulin because they directly explain diagnosis of diabetes.
```{r}
df<-df[,c(-2,-5)]
```

#Response and Primary Predictor Overview
```{r}
library(dplyr)
df %>% count(Outcome)
```

```{r}
hist(df$Pregnancies, xlab = 'Number of Pregnancies', main='Histogram of Pregnancies')
hist(df$Outcome,xlab='Outcome', main='Histogram of Outcome')
```


#Imputate mising data with mean:
```{r}
library(dplyr)
df %>% count(BloodPressure)
BP_mean = round(sum(df$BloodPressure)/(768-35),2)
df$BloodPressure[df$BloodPressure==0]<-BP_mean
```

```{r}
library(dplyr)
df %>% count(BMI)
BMI_mean = round(sum(df$BMI)/(768-11),2)
df$BMI[df$BMI==0]<-BMI_mean
```

#Imputate missing skinthickness data: 
```{r}
library(dplyr)
df %>% count(SkinThickness)
```

```{r}
df_comp <- df[,-7]
df_comp[df_comp==0]<-NA
df_comp$Outcome<-df$Outcome
df_comp = df_comp[complete.cases(df_comp),]
```

```{r,warning=FALSE}
model_st<-lm(SkinThickness ~ Age+BloodPressure+BMI+DiabetesPedigreeFunction, data=df_comp)
library(sjPlot)
tab_model(model_st, dv.labels = 'Imputation Model', digits = 3)
summary(model_st)
```

```{r}
ind = which(df$SkinThickness == 0) 
df$SkinThickness[ind] = predict(model_st, newdata = df[ind,])
df$SkinThickness<-round(df$SkinThickness,2)
```

#Data Preview after imputation
```{r}
library(psych)
data_view<-describeBy(df)
data_view<-data_view[, c(3,4,5,8,9,10)]
data_view
```

#Check correlations: 
```{r,warning=FALSE}
library(ggcorrplot)
df.cor<-round(cor(df),2)
ggcorrplot(df.cor)
```  

#Fit a model with confounder variables with backward selection: 
```{r}
model_conf1 <- glm(formula=Outcome ~ BloodPressure + SkinThickness + BMI + DiabetesPedigreeFunction + Age, family=binomial(link=logit), data=df)
library(sjPlot)
tab_model(model_conf1, dv.labels = 'Confounder Model', digits = 3)
step(model_conf1, scope=~1, direction="backward", p=0.2)
```

#Final confounder model: 
```{r}
model_conf2<-glm(formula = Outcome ~ BMI + DiabetesPedigreeFunction + Age, family = binomial(link = logit), data = df)
library(sjPlot)
tab_model(model_conf2, dv.labels = 'Final Confounder Model', digits = 3)
summary(model_conf2)
```

#Add main variable of interest - Pregnancies: 
```{r}
model_main<-glm(formula = Outcome ~ Pregnancies + BMI + DiabetesPedigreeFunction + Age, family = binomial(link = logit), data = df)
library(sjPlot)
tab_model(model_main, dv.labels = 'Model with Primary Interest', digits = 3)
summary(model_main)
```

#Interested in how age affects pregnancies' influence on diabetes
```{r}
model_main_int<-glm(formula = Outcome ~ Pregnancies + BMI + DiabetesPedigreeFunction + Age + Pregnancies:Age, family = binomial(link = logit), data = df)
anova(model_main, model_main_int, test='LRT')
```

```{r}
library(sjPlot)
tab_model(model_main_int, dv.labels = 'Model with Primary Interest and Interaction', digits = 3)
summary(model_main_int)
```

#Goodness of Fit: 
```{r,warning=FALSE}
hoslem=function (obs, pred, g = 0) {  if (g == 0){g = round(min(length(obs)/10, 10))}
ord <- order(pred)
obs.o <- obs[ord]
pred.o <- pred[ord]
interval = cut(pred.o, quantile(pred.o, 0:g/g), include.lowest = TRUE)
 counts = xtabs(formula = cbind(obs.o, pred.o) ~ interval)
 centers <- aggregate(formula = pred.o ~ interval, FUN = "mean")
 pear.res <- (counts[, "obs.o"] - counts[, "pred.o"])/sqrt(counts[, "pred.o"])
 pearson <- sum(pear.res^2)
 if (any(counts[, "pred.o"] < 5))
{ warning("Some expected counts are less than 5. Use smaller number of groups")}
 p = 1 - pchisq(pearson, g - 2)
 results <- as.data.frame(cbind(counts[, "obs.o"], counts[,"pred.o"], centers$pred.o, pear.res))
colnames(results) <- c("obs.o", "pred.o", "avg mean", "pearson resid")
cat("Hosmer-Lemeshow test with", g, "bins", "\n", "Pearson Stat = ",pearson, "\n", "p = ", p, "\n \n")

return(results)
}

HL.table = hoslem(df$Outcome, model_main_int$fitted.values)
```

#Diagnostics:
```{r}
par(mfrow=c(2,2))
plot(model_main_int)
```

```{r}
plot(model_main_int, 4)
```

```{r}
outlier<-rbind(df[59,], df[229,], df[488,])
outlier
```

#Final model of data with imputation:
```{r}
model_final<-glm(formula = Outcome ~ Pregnancies + BMI + DiabetesPedigreeFunction + Age + Pregnancies:Age, family = binomial(link = logit), data = df)
library(sjPlot)
tab_model(model_final, dv.labels = 'Final model with imputation', digits = 3)
```

```{r}
round(exp(model_final$coefficients),3)
```

```{r,warning=FALSE}
round(exp(confint(model_final)),3)
```


#Extract complete data
```{r}
df<-read.csv('diabetes.csv')
df<-df[,c(-2,-5)]
df_comp <- df[,-7]
df_comp[df_comp==0]<-NA
df_comp$Outcome<-df$Outcome
df_comp = df_comp[complete.cases(df_comp),]
```

#Fit a model with confounder variables with backward selection: 
```{r}
model2_conf1 <- glm(formula=Outcome ~ BloodPressure + SkinThickness + BMI + DiabetesPedigreeFunction + Age, family=binomial(link=logit), data=df_comp)
step(model2_conf1, scope=~1, direction="backward", p=0.2)
```

#Final confounder model: 
```{r}
model2_conf2<-glm(formula = Outcome ~ BMI + DiabetesPedigreeFunction + Age, family = binomial(link = logit), data = df_comp)
#summary(model2_conf2)
```

#Add main variable of interest - Pregnancies: 
```{r}
model2_main<-glm(formula = Outcome ~ Pregnancies + BMI + DiabetesPedigreeFunction + Age, family = binomial(link = logit), data = df_comp)
summary(model2_main)
```

#Add interaction: 
```{r}
model2_main_int<-glm(formula = Outcome ~ Pregnancies + BMI + DiabetesPedigreeFunction + Age + Pregnancies:Age, family = binomial(link = logit), data = df_comp)
anova(model2_main, model2_main_int, test='LRT')
```

#Diagnostics:
```{r}
par(mfrow=c(2,2))
plot(model2_main)
```

```{r}
plot(model2_main, 4)
```


```{r}
outlier<-rbind(df[229,], df[460,], df[674,])
outlier
```

#Goodness of Fit:
```{r,warning=FALSE}
HL.table = hoslem(df_comp$Outcome, model2_main_int$fitted.values)
```

#Final model of complete case data:
```{r}
model2_final<-glm(formula = Outcome ~ Pregnancies + BMI + DiabetesPedigreeFunction + Age + Pregnancies:Age, family = binomial(link = logit), data = df_comp)
```

#Sensitivity:
```{r}
library(sjPlot)
tab_model(model_final, model2_final, dv.labels = c("Final Model with Imputation", "Final Model with Completers"), digits = 3)
```



